### 2015-06-23 chapters/ipa_meets_unicode.tex
First draft of IPA recommendations

Summarizing the pitfalls as discussed in this chapter, we propose to define a 
``strict'' IPA encoding to be used when interoperability of phonetic resources 
is intended. For regular research papers, when only the visual outline is 
important, the details encoding of IPA in electronic format is rather 
unimportant, as long as it looks as expected. So for those regular usage these 
proposals are irrelevant.

TO DO: 
[different kind of profiles]

At the end of this chapter we have added a few longish tables summarizing all
159 different unicode codepoints that form the basis of strict-IPA encoding
(107 letters, 36 diacritics and 16 remaining symbols). Each of
these tables shows a typical glyph, and then lists the Unicode Codepoint,
Unicode Name and IPA description for each symbol.

TO DO:
[ordering]
[boundaries, stress, intonation]
[laxIPA]
[additions]

@@

  \item \textsc{IPA letters} \newline
        The 107 different IPA letters as allowed in strict-IPA encoding are
        listed in Table~\ref{tab:ipa_letters} starting on
        page~\pageref{tab:ipa_letters}.
  \item \textsc{IPA diacritics} \newline The 36 different IPA diacritics and
        tone markers (both Unicode Modifier Letters and Combining Diacritical
        Marks) as allowed in strict-IPA encoding are listed in
        Table~\ref{tab:ipa_diacritics} starting on
        page~\pageref{tab:ipa_diacritics}.
  \item \textsc{IPA remainders} \newline The 16 remaining IPA symbols
        (boundary, stress, tone letters and intonation markers) as allowed in strict-IPA
        encoding are listed in Table~\ref{tab:ipa_leftovers} on
        page~\pageref{tab:ipa_leftovers}.
  \item \textsc{IPA alternatives} \newline The 16 additional symbols as allowed in
        lax-IPA encoding are listed in Table~\ref{tab:ipa_lax} on
        page~\pageref{tab:ipa_lax}.     
  \item \textsc{optional IPA additions} \newline
        The 10 proposed additions to the IPA are listed in
        Table~\ref{tab:ipa_additions} on page~\pageref{tab:ipa_additions}.

@@

\newpage
\input{tables/unicode_letters}
\input{tables/unicode_diacritics}
\input{tables/unicode_boundaries}
\input{tables/unicode_lax}
\input{tables/unicode_additions}

>>>>

Summarizing the pitfalls as discussed in this chapter, we propose to define 
three different IPA encodings: strict-IPA, valid-IPA and widened-IPA.\@ These 
three encodings are subsets of each other, i.e.~strict-IPA is more restricted 
than valid-IPA, which it turn is more restricted than widened-IPA.\@ Informally 
speaking, valid-IPA represents the current state of the IPA \citep{IPA2005}. 
Strict-IPA represents a more constrained version of IPA, while widened-IPA is a 
slightly extended version of IPA, allowing a few more symbols.

\noindent Strict-IPA encoding is supposed to be used when interoperability of
phonetic resources is intended. It is a strongly constrained subset of IPA
geared towards uniqueness of encoding. Ideally, for each transcription there
should be exactly one possible strict-IPA encoding. For each phonetic feature
there is only one possibility (see Section~\ref{pitfall-multiple-options-ipa})
and the IPA diacritics are forced into a canonical ordering (see
Section~\ref{pitfall-no-unique-diacritic-ordering}).

Valid-IPA does allow alternative symbols with the same phonetic meaning, as 
specified in the official IPA specifications. Also, valid-IPA does not enforce a 
specific ordering of diacritics, because the IPA does not propose any such 
ordering. This means that in valid-IPA the same phonetic intention can be 
encoded in multiple ways. This is sufficient for phonetically trained human 
eyes, but it is not sufficient for automatic interoperability.

Finally, widened-IPA includes a few more symbols which seem to be useful for
various special cases (see Section~\ref{ipa-additions}).

\noindent At the end of this chapter we have added a few longish tables summarizing all
159 different unicode codepoints that form the basis of strict-IPA encoding
(107 letters, 36 diacritics and 16 remaining symbols). Each of
these tables shows a typical glyph, and then lists the Unicode Codepoint,
Unicode Name and IPA description for each symbol. Further, there is a table with 
the additional options for valid-IPA and a table with the additional options for 
widened-IPA.\@

textbf{NOTE: I HAVE ADDED VARIOUS TONE DIACRITICS TO VALID-IPA THAT ARE NOT IN
THE OFFICIAL IPA SUMMARY. SHOULD THEY BE INCLUDED?}

@@

  \item \textsc{strict-IPA letters} \newline
        The 107 different IPA letters as allowed in strict-IPA encoding are
        listed in Table~\ref{tab:ipa_letters} starting on
        page~\pageref{tab:ipa_letters}.
  \item \textsc{strict-IPA diacritics} \newline The 36 different IPA diacritics and
        tone markers (both Unicode Modifier Letters and Combining Diacritical
        Marks) as allowed in strict-IPA encoding are listed in
        Table~\ref{tab:ipa_diacritics} starting on
        page~\pageref{tab:ipa_diacritics}.
  \item \textsc{strict-IPA remainders} \newline The 16 remaining IPA symbols
        (boundary, stress, tone letters and intonation markers) as allowed in strict-IPA
        encoding are listed in Table~\ref{tab:ipa_leftovers} on
        page~\pageref{tab:ipa_leftovers}.
  \item \textsc{valid-IPA additions} \newline The 16 additional symbols as allowed in
        valid-IPA encoding are listed in Table~\ref{tab:ipa_lax} on
        page~\pageref{tab:ipa_lax}.     
  \item \textsc{widened-IPA additions} \newline
        The 10 proposed additions to the IPA are listed in
        Table~\ref{tab:ipa_additions} on page~\pageref{tab:ipa_additions}.

@@

\newpage
\input{tables/unicode_letters}
\input{tables/unicode_diacritics}
\input{tables/unicode_remaining}
\input{tables/unicode_valid}
\input{tables/unicode_additions}

### 2015-06-23 tables/unicode_valid.tex
First draft of IPA recommendations

\tablecaption{Alternatives within IPA with Unicode encodings}\label{tab:ipa_lax}
 \tablefirsthead{
   \toprule
   & Code & Unicode name & Phonetic description \\ 
   \midrule
   }
 \tablehead{
   \multicolumn{4}{c}{
     \captionsize\tablename\ \thetable{} 
     Alternatives within IPA with Unicode encodings --- \textit{continued}
     } \\
   \toprule
   & Code & Unicode name & Phonetic description \\ 
   \midrule
   }
 \tabletail{
   \bottomrule
   \multicolumn{4}{r}{
     \captionsize\textit{continued on next page}
     } \\
   }
 \tablelasttail{\bottomrule}

>>>>

\tablecaption{Additional characters for valid-IPA with Unicode encodings}\label{tab:ipa_lax}
 \tablefirsthead{
   \toprule
   & Code & Unicode name & Phonetic description \\ 
   \midrule
   }
 \tablehead{
   \multicolumn{4}{c}{
     \captionsize\tablename\ \thetable{} 
     Additional characters for valid-IPA with Unicode encodings --- \textit{continued}
     } \\
   \toprule
   & Code & Unicode name & Phonetic description \\ 
   \midrule
   }
 \tabletail{
   \bottomrule
   \multicolumn{4}{r}{
     \captionsize\textit{continued on next page}
     } \\
   }
 \tablelasttail{\bottomrule}
